---
title: "Calibrate CNN predictions for Rhizaria"
author: "Thelma Panaiotis"
date: "25/08/2022"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 3
    collapsed: false
    fig_width: 10
    fig_height: 8 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
#library(reticulate)
library(grid)
library(gridExtra)
library(scales)
library(arrow)
library(googlesheets4)
gs4_auth(use_oob = TRUE)
1

ss <- "https://docs.google.com/spreadsheets/d/100Hmfx9BVTFtBzw4pUETj6m3I4KjpibGvuRvDEVRYC8/edit?usp=sharing"

```

## Prepare data

Read predictions and scores of calibration data

```{r read_preds}
# Read taxo match to change EcoTaxa names to nice names
tc <- read_sheet(ss)

# True classification
df_true <- read_parquet("data/rhizaria_calibration_data.parquet") %>% 
  select(object_id = orig_id, classif_id_1) %>% 
  left_join(tc %>% select(classif_id_1, y_true = new_taxon)) %>% 
  select(-classif_id_1)

# Predictions
df_pred <- read_tsv("data/ecotaxa_rhizaria_calibration.tsv", comment = "[") %>% select(object_id, y_pred = object_annotation_category, score = object_cnn_score)

# Join truth and predictions
df <- left_join(df_true, df_pred)

```

Within each taxon, order by score and compute cumulative error as well as rank.

```{r rank}
calib_preds <- df %>% 
  # Group by taxon
  group_by(y_pred) %>% 
  # Sort by score
  arrange(score) %>% 
  # Compute cumulative error in % and rank in % 
  mutate(
    error = y_true != y_pred,
    cum_error = cumsum(error),
    nb_error = max(cum_error),
    cum_error = ifelse(nb_error>0, cum_error/nb_error, 1), # Treat cases of zero division when all objects are correctly predicted
    percent_rank=1:n()/n()
  ) %>% 
  ungroup() %>% 
  # Count objects per true class
  add_count(y_true) %>% 
  select(-nb_error)
```


## Inspect data

Count objects in each true and predicted class and plot them

```{r counts_y_true}
counts_y_true <- calib_preds %>% 
  count(y_true) %>% 
  arrange(desc(n)) %>% 
  mutate(y_true = fct_inorder(y_true))

counts_y_true %>% 
  ggplot() +
  geom_col(aes(x = y_true, y = n, fill = n > 10)) +
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Composition of calibration set")
```

*`r sum(counts_y_true$n < 10)` classes have fewer than 10 objects: computing a threshold for theses classes will be difficult.*

```{r counts_y_pred}
counts_y_pred <- calib_preds %>% 
  count(y_pred) %>% 
  arrange(desc(n)) %>% 
  mutate(y_pred = fct_inorder(y_pred))
  
counts_y_pred %>% 
  ggplot() +
  geom_col(aes(x = y_pred, y = n, fill = n > 10)) +
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Predictions in calibration set")
```

*More objects are predicted in plankton classes than the ones in true classes: the classifier has a good recall, and the calibration step will improve precision.*

Now look at scores distribution for each taxon. 

```{r scores_dist}
calib_preds %>% 
  ggplot() +
  geom_density(aes(x = score)) +
  facet_wrap(~y_pred, scales = "free_y") +
  labs(title = "Score distribution for predicted classes")
```


## Global view for thresholding

For each taxon plot cumulative error VS prediction score. 

```{r score_distribution}
calib_preds %>% 
  ggplot() +
  geom_density(aes(x = score)) +
  facet_wrap(~y_pred, scales = "free_y") +
  labs(title = "Score distribution")
```

```{r cum_error_score}
calib_preds %>% 
  ggplot() +
  geom_path(aes(x = score, y = cum_error)) +
  facet_wrap(~y_pred) +
  scale_color_continuous(trans = "log1p") +
  labs(title = "Cumulative error VS prediction scores for pred classes")
```

And plot cumulative error VS object rank. 

```{r cum_error_rank}
calib_preds %>% 
  ggplot() +
  geom_path(aes(x = percent_rank, y = cum_error)) +
  facet_wrap(~y_pred) +
  scale_color_continuous(trans = "log1p") +
  labs(title = "Cumulative error VS prediction rank for pred classes")
```

For many taxa, cutting at a cumulative error of 90% will keep the major parts of the objects. Let’s now compute a threshold score for each taxon and see what proportion of data will be kept. 


## Thresholding by taxon

Make a list of taxa.
```{r taxa}
taxa <- calib_preds %>% pull(y_pred) %>% unique() %>% sort()
taxa
```

There are `r length(taxa)` taxa to compute threshold for.

Define a base error threshold and prepare a tibble to store computed thresholds. 

```{r thresholds_df}
error_threshold_base <- 0.9

taxa_threshold <- tibble(
  taxa, 
  error_threshold = NA,
  score_threshold = NA,
  keep_threshold = NA,
  )

```

For each taxon, start with an error threshold of 90%. Plot cumulative error VS score and VS rank and compute the score threshold associated with this error threshold and the proportion of objects that will be kept. 


```{r}
plot_threshold <- function(taxon, error_threshold = 0.9){
  #' Plot cumulative error VS score and VS rank for a given error threshold.
  #' Return the score threshold associated with this error threshold 
  #' and the proportion of objects that will be kept.  
  
  calib_taxa <- calib_preds %>% 
    filter(y_pred == taxon) %>% 
    mutate(keep = cum_error >= error_threshold)
  
  # Compute associated score threshold and proportion of kept data. 
  score_threshold <- calib_taxa %>% filter(keep) %>% slice(1) %>% pull(score)
  keep_threshold <- calib_taxa %>% filter(keep) %>% slice(1) %>% pull(percent_rank)

  # If error_threshold was set to 0, keep all objects
  if (error_threshold == 0) {
    score_threshold = 0
    keep_threshold = 0
  }
  
  # Number of true objects for this taxon
  n <- nrow(calib_taxa)
    
  
  if (n > 20000){ # if too many objects, do not plot 
    plot <- NULL
  } else {  
    p1 <- calib_taxa %>% 
      ggplot() +
      geom_vline(aes(xintercept = score_threshold)) +
      geom_label(aes(x = 1, y = 0, label = format(score_threshold, digits = 2)))
    
    p2 <- calib_taxa %>% 
      ggplot() +
      geom_vline(aes(xintercept = keep_threshold)) + 
      geom_label(aes(x = 1, y = 0, label = str_c("Keep ", percent(1-keep_threshold, digits = 2), " of objects")))
  
    if (n < 1000){ # for few objects, plot points
      p1 <- p1 + geom_point(aes(x = score, y = cum_error, colour = cum_error < error_threshold), show.legend = F)
      p2 <- p2 + geom_point(aes(x = percent_rank, y = cum_error, colour = cum_error < error_threshold), show.legend = F)
    } else { # else, plot a path
      p1 <- p1 + geom_path(aes(x = score, y = cum_error, colour = cum_error < error_threshold), show.legend = F)
      p2 <- p2 + geom_path(aes(x = percent_rank, y = cum_error, colour = cum_error < error_threshold), show.legend = F)
    }
  
    # Make 2 plots side by side
    plot <- grid.arrange(p1, p2, ncol=2, top = textGrob(paste0(taxon, " (n = ", n, ")"), gp=gpar(fontsize=18)))  
  }
  return(list(plot, score_threshold, keep_threshold))
}
```


### Acantharea
```{r Acantharea, echo=F}
taxon <- "Acantharea"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

Ok for error threshold at 90%.

### Acantharea_like

```{r Acantharea_like, echo=F}
taxon <- "Acantharea_like"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

### Acantharea_small

```{r Acantharea_small, echo=F}
taxon <- "Acantharea_small"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon, error_threshold = error_threshold)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```


### Arthracanthida

```{r Arthracanthida, echo=F}
taxon <- "Arthracanthida"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon, error_threshold = error_threshold)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```


### Aulacanthidae

```{r Aulacanthidae, echo=F}
taxon <- "Aulacanthidae"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

### Aulacanthidae_flat

```{r Aulacanthidae_flat, echo=F}
taxon <- "Aulancanthidae_flat"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```


### Aulosphaeridae

```{r Aulosphaeridae, echo=F}
taxon <- "Aulosphaeridae"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```
Ignore these.

### Collodaria_colonial

```{r Collodaria_colonial, echo=F}
taxon <- "Collodaria_colonial"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon, error_threshold = error_threshold)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

### Collodaria_like

```{r Collodaria_like, echo=F}
taxon <- "Collodaria_like"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

Ok for error threshold at 90%.


### Collodaria_solitaryblack

```{r Collodaria_solitaryblack, echo=F}
taxon <- "Collodaria_solitaryblack"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```



### Collodaria_solitaryglobule

```{r Collodaria_solitaryglobule, echo=F}
taxon <- "Collodaria_solitaryglobule"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```



### Foraminifera

```{r Foraminifera, echo=F}
taxon <- "Foraminifera"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon, error_threshold = error_threshold)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```





### Spumellaria

```{r Spumellaria, echo=F}
taxon <- "Spumellaria"
index <- which(taxa == taxon)
error_threshold <- 0.9

# Compute
thr <- plot_threshold(taxon = taxon)

# Fill the threshold dataframe
taxa_threshold$error_threshold[index] <- error_threshold
taxa_threshold$score_threshold[index] <- thr[[2]]
taxa_threshold$keep_threshold[index] <- thr[[3]]
```

Ok for error threshold at 90%.

## Thresholding summary 
```{r summary}
taxa_threshold <- taxa_threshold %>% 
  #mutate(plankton = taxa %in% calibration_data$plankton_classes) %>% 
  mutate(retained = 1 - keep_threshold) %>% 
  mutate(usable = retained > 0.1)

write_csv(taxa_threshold, file = "data_git/07.taxa_threshold_rhizaria.csv")

taxa_threshold %>% 
  arrange(-retained) %>% 
  mutate(taxa = factor(taxa, levels = taxa)) %>% 
  ggplot() +
  geom_col(aes(x = taxa, y = retained)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Proportion of retained objects")
```

